---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ZIPG

<!-- badges: start -->
<!-- badges: end -->

We provide R code for Zero-inflated Poisson-Gamma Model (ZIPG) with an application to longitudinal microbiome count data. 

## Installation

You can install the development version of ZIPG like so:

``` r
devtools::install_github("roulan2000/ZIPG")
```

## Example

### Load Data

Complete Dietary data can be find in "Daily sampling reveals personalized diet-microbiome associations in humans." (Johnson et al. 2019)

```{r}
library(ZIPG)
library(ggplot2)
data("Dietary")
dat = Dietary
taxa_num = 100
dat$taxa_name[taxa_num] # taxa name
W = dat$OTU[,taxa_num] # taxa count
M = dat$M # sequencing depth
ggplot(NULL)+
  geom_boxplot(aes(
  x = as.factor(dat$COV$ALC01),y=log((W+0.5)/M)))+
  labs(title = dat$taxa_name[taxa_num],
       x = 'ALC',y='Relative Abundance')+
  theme_bw()
```


## ZIPG Wald

Use function `ZIPG_main()` to run our ZIPG model.

**Input :**

`W` : Observed taxa count data.

`M` : Sequencing depth, ZIPG use log(M) as offset by default.

`X`, `X_star` : Covariates of interesting of differential abundance and differential varibility, input as formula.

**Output list:**

`ZIPG_res$init` : pscl results, used as initialization.

`ZIPG_res$res` : ZIPG output evaluated at last EM iteration.

`ZIPG_res$res$par` : ZIPG estimation for $\Omega = (\beta,\beta^*,\gamma)$.
  
`ZIPG_res$wald_test` : ZIPG Wald test

`ZIPG_res$logli` : ZIPG log-likelihood

```{r}
ZIPG_res <- ZIPG_main(data = dat$COV,
                      X = ~ALC01+nutrPC1+nutrPC2, X_star = ~ ALC01,
                      W = W, M = M)
res = ZIPG_summary(ZIPG_res)
```

## ZIPG bWald

Set the bootstrap replicates `B` in `bWald_list` to conduct ZIPG-bWald, results and covariance matrix can be find in `ZIPG_res$bWald`. 

```{r}
# Set bootstrap replicates B
bWald_list = list(B = 100)
# Wait for a wile
ZIPG_res1 = ZIPG_main(
  data = dat$COV,
  X = ~ALC01+nutrPC1+nutrPC2, X_star = ~ ALC01,
  W = W, M = M,
  bWald_list = bWald_list)
res = ZIPG_summary(ZIPG_res1,type = 'bWald')
res = ZIPG_CI(ZIPG_res1)
```

To test more complicated hypothesis, you may use the covariance matirx driven from bootstrap.

```{r}
round(ZIPG_res1$bWald$vcov,3)
```

## ZIPG pbWald

Set bootstrap replicates `B` and the null hypothesis by formula `X0` and `X_star0` in `pbWald_list` to conduct ZIPG-pbWald, results can be find in ZIPG_res\$pbWald

```{r}
# test beta1star, the 6th parameter
# 
pbWald_list = list(
  X0 = ~ALC01 + nutrPC1+nutrPC2,
  X_star0 = ~ 1,
  B = 100
)

ZIPG_res2 = ZIPG_main(
  data = dat$COV,
  X = ~ALC01+nutrPC1+nutrPC2, X_star = ~ ALC01,
  W = W, M = M,
  pbWald_list= pbWald_list)

res = ZIPG_summary(ZIPG_res2,type ='pbWald')
```
